Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-05-10 apj.tex
housekeeping

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/ebv.pdf}
   \includegraphics[width=2.8in]{output23/ebv_delta.pdf}
      \caption{Left: Ideograms of the external $E(B-V)$ and
   internal $E_\delta(B-V) = (\delta_B-\delta_V)D$  contributions to color excess  for the supernovae in our sample.
   Right: Ideogram and histogram of the medians of $E_\delta(B-V)$ on an enlarged scale.
   \label{ebv:fig}}
\end{figure}
The significance of the $\delta$ parameters is shown pictorially in
Figure~\ref{deltacorner:fig}, which shows the $\delta$ posteriors.
Recall that fixing $\delta_X=0$ reverts this model into the two extrinsic-parameter model of \S\ref{model:sec}.
Assigning the Monte Carlo links to bins of width 0.5 in each $\delta$ parameter, the maximum number of links in a single bin is 303, the number of bins
with at least one element is 1987, and the bin that contains $\delta=0$ for all bands has zero elements; we thus claim a detection
of a third color-parameter at $> (1-5\times 10^{-4})$.
Figure~\ref{deltaratio:fig} shows  values of $\delta_X/\delta_I-1$ that  have significant non-zero values.
While the signal of the internal parameter is strong within the context of this model, ``model testing'' (comparing the
performance  between this model
and that of \S\ref{model:sec}), is challenging for Bayesian Monte Carlo analysis.  Evidence calculations are planned
for the next generation of STAN and the development of robust evidence algorithms is a subject of theoretical work.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23/delta_corner.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the posterior contours for $\delta$ for Model~II.
   Lines for $\delta=0$ are overplotted.
   \label{deltacorner:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   A dotted line at zero is shown for reference.
   \label{deltaratio:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, $E_{\gamma^1}(B-V)$,  $A_{\delta I}=\delta_I D$ are given in the matrix
\color{purple}
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V),  A_{\delta I}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrr}
1& 0.00^{+0.06}_{-0.06} & -0.07^{+0.06}_{-0.07} & -0.04^{+0.06}_{-0.06} & 0.11^{+0.07}_{-0.07} & -0.00^{+0.07}_{-0.07} & -0.06^{+0.07}_{-0.06} \\
0.00^{+0.06}_{-0.06} & 1 & 0.11^{+0.03}_{-0.04} & -0.26^{+0.03}_{-0.03} & -0.07^{+0.06}_{-0.06} & -0.01^{+0.08}_{-0.08} & -0.01^{+0.08}_{-0.08} \\
-0.07^{+0.06}_{-0.07} & 0.11^{+0.03}_{-0.04} & 1 & -0.14^{+0.04}_{-0.03} & -0.17^{+0.05}_{-0.05} & -0.05^{+0.08}_{-0.07} & 0.04^{+0.07}_{-0.07} \\
-0.04^{+0.06}_{-0.06} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.04}_{-0.03} & 1 & 0.03^{+0.06}_{-0.06} & 0.04^{+0.07}_{-0.08} & -0.08^{+0.10}_{-0.09} \\
0.11^{+0.07}_{-0.07} & -0.07^{+0.06}_{-0.06} & -0.17^{+0.05}_{-0.05} & 0.03^{+0.06}_{-0.06} & 1 & 0.06^{+0.08}_{-0.07} & 0.06^{+0.08}_{-0.07} \\
-0.00^{+0.07}_{-0.07} & -0.01^{+0.08}_{-0.08} & -0.05^{+0.08}_{-0.07} & 0.04^{+0.07}_{-0.08} & 0.06^{+0.08}_{-0.07} & 1 & 0.01^{+0.08}_{-0.07} \\
-0.06^{+0.07}_{-0.06} & -0.01^{+0.08}_{-0.08} & 0.04^{+0.07}_{-0.07} & -0.08^{+0.10}_{-0.09} & 0.06^{+0.08}_{-0.07} & 0.01^{+0.08}_{-0.07} & 1 \\
\end{array}
\end{pmatrix}.
\end{multline}
\color{black}
The correlation coefficients between the ``intrinsic'' and ``extrinsic'' parameters are
\color{purple}
not statistically significant.
\color{black}
This has two important implications:
First, while in principle a single physical color effect can affect both intrinsic and extrinsic parameters, from the correlations
there is no evidence
that this occurs.
Second, the intrinsic term is not a second-order extrinsic correction.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23/perobject_corner.pdf} 
   \caption{
   Distributions for the supernova parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$,  $E_{\gamma^1}(B-V)$,  and $E_{\delta}(B-V)$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject2:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23_sim/coeff0.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $U$ band of ten independent simulated data sets.  Lines show the values of the input values.
            \label{simglobal1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23_sim/coeff1.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $B$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal2:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23_sim/coeff2.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $V$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal3:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23_sim/coeff3.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $R$ band of ten independent simulated data sets.   Lines show the values of the input values.
\label{simglobal4:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
         \includegraphics[width=5.2in]{output23_sim/coeff4.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $I$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal5:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
%   \includegraphics[width=2.8in]{output23/x1si.pdf}
   \includegraphics[width=5in]{output23/x1D.pdf}
    \caption{The SALT2 $x_1$ versus $A_{\delta I}  = \delta_I D$.
   \label{x1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23/residualx1.pdf}
    \caption{Differences between observed colors and the colors predicted from the analysis, as a function
            of the SALT2 light-curve shape parameter $x_1$.  Our model does not explicitly use $x_1$.  For reference a dotted line is plotted at zero difference.
   \label{x1res:fig}}
\end{figure}

@@

\subsection{Host-Galaxy Mass}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that when introducing intrinsic color scatter as a latent parameter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore whether the Hubble residual may be due to unaccounted color,
we plot in Figure~\ref{childress:fig} our intrinsic parameter  $A_{\delta I}$  versus host mass 
\color{purple}
for the subset of supernovae whose host measurements are given in \citet{2013ApJ...770..108C}.
Also plotted  are ideograms of $A_{\delta I}$ for the two subsamples obtained by dividing the full sample at a host mass of  $\log{(M/M_\sun)}=10$.
Although by eye it appears that the high-mass hosts tend the have lower $A_{\delta I}$, the difference is
not significant; we calculate  average
%---
$\langle A_{\delta I} \rangle=   -0.0026 \pm    0.0014$,
$\langle A_{\delta I} \rangle=-0.0035 \pm    0.0010$ 
for low- and high-mass hosts respectively. 
%Using the SNfactory sample,
%\citet{2013A&A...560A..66R,2017Rigault} find that a step in Hubble residuals is better related to local star formation rate, rather than
%global host mass.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/childress.pdf}
   \includegraphics[width=2.8in]{output23/childress2.pdf}
      \caption{Left: Intrinsic parameter $A_{\delta I}=\delta_I D$  versus host galaxy mass. Overplotted are the mean and 1$\sigma$ values for supernovae with hosts
      less than and greater than  $\log{(M/M_\sun)}=10$.
Right: Normalized ideograms of  $A_{\delta I}$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
   \label{childress:fig}}
\end{figure}

>>>>

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output25/ebv.pdf}
   \includegraphics[width=2.8in]{output25/ebv_delta.pdf}
      \caption{Left: Ideograms of the external $E(B-V)$ and
   internal $E_\delta(B-V) = (\delta_B-\delta_V)D$  contributions to color excess  for the supernovae in our sample.
   Right: Ideogram and histogram of the medians of $E_\delta(B-V)$ on an enlarged scale.
   \label{ebv:fig}}
\end{figure}
The significance of the $\delta$ parameters is shown pictorially in
Figure~\ref{deltacorner:fig}, which shows the $\delta$ posteriors.
Recall that fixing $\delta_X=0$ reverts this model into the two extrinsic-parameter model of \S\ref{model:sec}.
Assigning the Monte Carlo links to bins of width 0.5 in each $\delta$ parameter, the maximum number of links in a single bin is 303, the number of bins
with at least one element is 1987, and the bin that contains $\delta=0$ for all bands has zero elements; we thus claim a detection
of a third color-parameter at $> (1-5\times 10^{-4})$.
Figure~\ref{deltaratio:fig} shows  values of $\delta_X/\delta_I-1$ that  have significant non-zero values.
While the signal of the internal parameter is strong within the context of this model, ``model testing'' (comparing the
performance  between this model
and that of \S\ref{model:sec}), is challenging for Bayesian Monte Carlo analysis.  Evidence calculations are planned
for the next generation of STAN and the development of robust evidence algorithms is a subject of theoretical work.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25/delta_corner.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the posterior contours for $\delta$ for Model~II.
   Lines for $\delta=0$ are overplotted.
   \label{deltacorner:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output25/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   A dotted line at zero is shown for reference.
   \label{deltaratio:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, $E_{\gamma^1}(B-V)$,  $A_{\delta I}=\delta_I D$ are given in the matrix
\color{purple}
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V),  A_{\delta U}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrr}
1 & 0.01^{+0.06}_{-0.06} & -0.07^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.07} & 0.10^{+0.07}_{-0.06} & 0.00^{+0.08}_{-0.07} & -0.01^{+0.08}_{-0.08} \\
0.01^{+0.06}_{-0.06} & 1 & 0.10^{+0.04}_{-0.04} & -0.26^{+0.03}_{-0.03} & -0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} & -0.00^{+0.08}_{-0.08} \\
-0.07^{+0.07}_{-0.06} & 0.10^{+0.04}_{-0.04} & 1 & -0.14^{+0.03}_{-0.03} & -0.16^{+0.05}_{-0.05} & -0.03^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} \\
-0.04^{+0.06}_{-0.07} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.03}_{-0.03} & 1 & 0.03^{+0.05}_{-0.06} & 0.02^{+0.07}_{-0.08} & -0.01^{+0.08}_{-0.08} \\
0.10^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.06} & -0.16^{+0.05}_{-0.05} & 0.03^{+0.05}_{-0.06} & 1 & 0.05^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} \\
0.00^{+0.08}_{-0.07} & -0.00^{+0.07}_{-0.08} & -0.03^{+0.08}_{-0.08} & 0.02^{+0.07}_{-0.08} & 0.05^{+0.08}_{-0.08} & 1 & -0.00^{+0.07}_{-0.08} \\
-0.01^{+0.08}_{-0.08} & -0.00^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} & -0.01^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} & -0.00^{+0.07}_{-0.08} & 1 \\
\end{array}
\end{pmatrix}.
\end{multline}
\color{black}
The correlation coefficients between the ``intrinsic'' and ``extrinsic'' parameters are
\color{purple}
not statistically significant.
\color{black}
This has two important implications:
First, while in principle a single physical color effect can affect both intrinsic and extrinsic parameters, from the correlations
there is no evidence
that this occurs.
Second, the intrinsic term is not a second-order extrinsic correction.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25/perobject_corner.pdf} 
   \caption{
   Distributions for the supernova parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$,  $E_{\gamma^1}(B-V)$,  and $E_{\delta}(B-V)$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject2:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25_sim/coeff0.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $U$ band of ten independent simulated data sets.  Lines show the values of the input values.
            \label{simglobal1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25_sim/coeff1.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $B$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal2:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25_sim/coeff2.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $V$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal3:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output25_sim/coeff3.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $R$ band of ten independent simulated data sets.   Lines show the values of the input values.
\label{simglobal4:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
         \includegraphics[width=5.2in]{output25_sim/coeff4.pdf} 
            \caption{Aggregated Model~II posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$,  $\delta$, and $\sigma$ in the $I$ band of ten independent simulated data sets.   Lines show the values of the input values.
 \label{simglobal5:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
%   \includegraphics[width=2.8in]{output25/x1si.pdf}
   \includegraphics[width=5in]{output25/x1D.pdf}
    \caption{The SALT2 $x_1$ versus $A_{\delta I}  = \delta_I D$.
   \label{x1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25/residualx1.pdf}
    \caption{Differences between observed colors and the colors predicted from the analysis, as a function
            of the SALT2 light-curve shape parameter $x_1$.  Our model does not explicitly use $x_1$.  For reference a dotted line is plotted at zero difference.
   \label{x1res:fig}}
\end{figure}

@@

\subsection{Host-Galaxy Mass}
A correlation between Hubble residual and host-galaxy mass
was first noted by \citet{2010ApJ...715..743K,2010MNRAS.406..782S}, a signal confirmed to exist in the SNfactory
sample \citep{2013ApJ...770..108C}.
This host-mass bias could be the result of a parameter that was not accounted for in the inference of SN~Ia absolute magnitude.
\citet{2016arXiv160904470M} find that when introducing intrinsic color scatter as a latent parameter, the null mass-step effect falls on the 95\%-level of their posterior.
To explore whether the Hubble residual may be due to unaccounted color,
we plot in Figure~\ref{childress:fig} our intrinsic parameter  $A_{\delta I}$  versus host mass 
\color{purple}
for the subset of supernovae whose host measurements are given in \citet{2013ApJ...770..108C}.
Also plotted  are ideograms of $A_{\delta U}$ for the two subsamples obtained by dividing the full sample at a host mass of  $\log{(M/M_\sun)}=10$.
Although by eye it appears that the high-mass hosts tend the have lower $A_{\delta U}$, the difference is
not significant; we calculate  average
%---
$\langle A_{\delta I} \rangle=   -0.0021 \pm    0.007$,
$\langle A_{\delta I} \rangle= -0.0034 \pm    0.0064$ 
for low- and high-mass hosts respectively. 
%Using the SNfactory sample,
%\citet{2013A&A...560A..66R,2017Rigault} find that a step in Hubble residuals is better related to local star formation rate, rather than
%global host mass.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output25/childress.pdf}
   \includegraphics[width=2.8in]{output25/childress2.pdf}
      \caption{Left: Intrinsic parameter $A_{\delta U}=\delta_U D$  versus host galaxy mass. Overplotted are the mean and 1$\sigma$ values for supernovae with hosts
      less than and greater than  $\log{(M/M_\sun)}=10$.
Right: Normalized ideograms of  $A_{\delta U}$ for supernovae in $\log{(M/M_\sun)}<10$ and $\log{(M/M_\sun)}>10$ hosts. 
   \label{childress:fig}}
\end{figure}
